Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in March - October 2020: An exploratory analysis

### 2021-06-14 manuscript/manuscript.Rmd
Changed formatting to rticles::plos

---
title: "Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in March - October 2020: An exploratory analysis" 
output: 
  #rticles::arxiv_article
  #pdf_document
  # bookdown::pdf_document2:
  #    toc: false
  bookdown::word_document2: 
    toc: false
    number_sections: false
    reference_docx: "output_template.docx"
abstract: |  
  **Abstract:** We collected a multi-centric retrospective dataset of patients (N = 213) who were admitted to ten hospitals in Czech Republic and tested positive for SARS-CoV-2 during the early phases of the pandemic in March - October 2020. The dataset contains baseline patient characteristics, breathing support required, pharmacological treatment received and multiple markers on daily resolution. Patients in the dataset were treated with hydroxychloroquine (N = 108), azithromycin (N = 72), favipiravir (N = 9), convalescent plasma (N = 7), dexamethasone (N = 4) and remdesivir (N = 3), often in combination. To explore association between treatments and patient outcomes we performed multiverse analysis, observing how the conclusions change between defensible choices of statistical model, predictors included in the model and other analytical degrees of freedom.  Weak evidence to constrain the potential efficacy of azithromycin and favipiravir can be extracted from the data. Additionally, we performed external validation of several proposed prognostic models for Covid-19 severity showing that they mostly perform unsatisfactorily on our dataset.
bibliography: Covid19.json
csl: journal-of-medical-virology.csl
#date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
author: "`r { source(here::here('manuscript', 'authors.R')); get_authors_string() }`"
---

```{r setup, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8)

@@

A convenience sample of patients from `r n_hospitals` sites was collected. The study sites span the whole spectrum of sizes from large university hospitals in major cities with multiple dedicated Covid-19 wards (Thomayer University Hospital in Prague, Motol University Hospital in Prague,  Kralovské Vinohrady University Hospital in Prague, General University Hospital in Prague, University Hospital in Pilsen) through major regional/specialized hospitals  (Na Homolce Hospital in Prague, Military Hospital Olomouc) as well as smaller hospitals caring for just several Covid-19 patients at a time (AGEL Hospital Nový Jičín, Hořovice Hospital, Třebíč Hospital). The sites were chosen based on availability and willingness of the personnel to participate in data collection. None of the study sites was exclusively dedicated to treating Covid-19 patients. For each site, the dataset contains all patients hospitalized in the participating wards over the data collection period. The data collection started at the onset of the Covid-19 pandemic in March 2020 (except for one site where some older records were inaccessible), but the end date for collection differs between sites due to time constraints of the participating physicians. Three sites included total of `r n_second_wave` patients that could be considered part of "second wave" (admitted after September 1st). Last patient included in the dataset was admitted on October 12th. See Supplementary Figure 1 for per-site data collection periods. Patients over the age of 18 were included if they had PCR-confirmed infection of SARS-CoV-2 and were not participating in a clinical trial of any Covid-19 pharmacotherapy.

@@

We collected data on comorbidities and information about disease progression on daily resolution including breathing support required, oxygen flow rate, experimental anti-Covid-19 and antimicrobial drugs taken and several laboratory markers (PCR positivity for SARS-CoV-2, C-reactive protein, D-dimer, Interleukin 6, Ferritin, lymphocyte count). Full protocol for data collection is attached in the supplementary material. Due to very low number of patients using extra-corporeal membrane oxygenation (N = `r n_ecmo`) or non-invasive positive pressure ventilation (N = `r n_nippv`) in our sample, we merged those categories with mechanical ventilation.

@@

First class of models are frequentist survival and multistate models under the proportional hazards assumption as implemented in the `coxph` function from the `survival` package [@http://zotero.org/users/5567156/items/I9UVTUJF]. We primarily use a model with competing risks for death and discharge from hospital (see Figure&nbsp;\@ref(fig:modelstates)a).

Second class of models are Bayesian hidden Markov models (HMM) of disease progression implemented via a custom extension to the `brms` package [@http://zotero.org/users/5567156/items/7ER3ZSER]. The parametrization of the HMM is inspired by Williams et al.[@http://zotero.org/users/5567156/items/F9K2L5Q8]: the actual process of disease is assumed to be continuous and allow only for transitions between neighboring states (as shown in Figure&nbsp;\@ref(fig:modelstates)b, c). The total probability of transition between any two states over the period of a day is then computed as the total probability of transition across all possible paths. This class of models does not satisfy the proportional hazards assumption, instead, it is assumed the process has the Markov property - i.e., that the (potentially unobserved) state and the covariates at a given day contain all the information to determine probabilities of the states on the next day. We use two versions of such models, one working solely with the observed breathing support and one assuming a hidden improving/worsening distinction. All of the hidden Markov models take into account whether best supportive care was initiated and a patient was thus not indicated to progress to more intensive treatment modalities.

@@

Except for age, sex and comorbidities, all covariates are treated as time-varying, e.g., the effect of taking a drug is only included for the days after the drug was taken. More details on the exact model formulations can be found in the supplementary statistical analysis.

@@

In total, we were able to gather data for `r n_patients` patients, see Table \@ref(tab:summarytable) for the overall characteristics of the patient sample and several subgroups we used in the analysis, including treatments taken. 
Counts of all treatment combinations are shown in Supplementary Figure 2 and Supplementary Figure 3 shows outcomes by study site, demonstrating quite large hospital-specific differences.
The dataset includes 19 patients already reported in a study of inflammatory signatures of Covid-19 [@http://zotero.org/users/5567156/items/IA4FQCF3].

@@

In Figure&nbsp;\@ref(fig:progression) we show the overall disease progression for all patients and in Figure&nbsp;\@ref(fig:selectedmarkers) we show the time-course of a subset of the markers we have measured. The data show some interesting patterns: patients with low Interleukin-6 or D-dimer values are overrepresented among patients with better outcomes, most patients had high CRP upon admission and for many the CRP levels stayed elevated over the whole hospitalization. However, the limited nature of the data does not allow for any statistically robust conclusions. We also see that the marker levels were not substantially stratified by study site.  Those patterns should however be interpreted with care due to systematic missingness of the data - in particular, patients that fared worse were probably more likely to have the markers measured. However, we believe this kind of patient-level view is useful to appreciate the extent of both between-patient and within-patient variability.

The between-patient variability is notable even across outcomes - when ordering the patients by the highest CRP levels experienced throughout the hospital stay, the top 20% of patients that breathed ambient air for the whole hospitalization experienced higher levels than the bottom 20% of patients that required ventilation or died. This overlap is even larger when comparing only against the patients that died and D-dimer, Interleukin-6 and lymphocyte count also show a notably larger overlap than CRP (Supplementary Figure 4).

@@

```{r selectedmarkers, fig.cap = "Values of selected markers over the course of the disease. Each line represents a patient, stratified by the worst breathing support required over the hospitalization. Color indicates study sites. The vertical scale is logarithmic. Ventilated = any form of ventilation (non-invasive positive-pressure ventilation, mechanical ventilation and extra-corporeal membrane oxygenation), CRP = C-reactive protein [mg/l], D-dimer [ng/ml DDU], Ly = lymphocyte count [10^9/l], IL-6 = Interleukin 6 [ng/l]."}
#data$marker_data %>%
# TODO: use SpO2 in AA as SpO2 native (and vice versa)
plot_marker_data <- function(markers, log_transform, smooth = FALSE, max_patients = NULL, max_day = NULL) {

@@

As noted above, the nature of the convenience sample did not enforce random assignment of treatments to patients. In fact, patients with worse baseline characteristics, which lead to worse outcomes, were less likely to receive hydroxychloroquine (see Figure&nbsp;\@ref(fig:hcqassoc)). This clearly creates a bias towards a positive effect of hydroxychloroquine on the outcome (and potentially for other treatments as well - most were used in combination with hydroxychloroquine), which, however, could be false.

```{r hcqassoc, message=FALSE, fig.width=7, fig.height=4.5, fig.cap="The choice of treatment with hydroxychloroquine seemed to be associated with the status of patients upon hospitalization. Comorbidities were associated with both worse outcome (black) and lower chance of treatment with hydroxychloroquine (red). Dots and lines represent the estimates and the 95% confidence intervals of the log odds ratio of the respective outcome. HCQ: hydroxychloroquine, IHD: ischemic heart disease, HD: hypertension drugs, HF: heart failure history, COPD: chronic obstructive pulmonary disease, LungD: other lung disease, Dia: diabetes, RenalD: renal disease, LiverD: liver disease, HighCr: creatinin above 115 for males or above 97 for females, HighInr: Prothrombin time (Quick test) as International Normalized Ratio above 1.2, LowAlb: albumin in serum/plasma below 36 g/l."}

@@

Here, we focus on hydroxychloroquine and azithromycin as those are the only treatments with larger sample size. We also investigate favipiravir as it is less well reported in the literature. Hydroxychloroquine was dosed almost exclusively in a 5-day regime starting with a loading dose of 800mg on the first day and followed by 400mg. Majority of patients complemented hydroxychloroquine with azithromycin while azithromycin was rarely used alone (see Table \@ref(tab:summarytable)). Azithromycin was most frequently dosed 250 or 500mg/day, but doses ranging from 100mg/day to 1500 mg/day were observed.  Favipiravir was used only at one site with a loading dose of 3600mg on the first day, followed by at most 9 days with a 1600mg dose. All but one of the patients receiving favipiravir also received hydroxychloroquine. Treatment was initiated mostly within two days of admission (see Supplementary Figure 5).

The results of the multiverse analysis for association between hydroxychloroquine, azithromycin and favipiravir usage and death is shown in Figure&nbsp;\@ref(fig:multiverseresultsdrugs) - here, we only show models that were not found to have immediate problems representing the data well or computational issues (see Supplementary statistical analysis for details). Results for all models we tested are reported in  supplementary Figures `r supplementary_figures_multiverse`, with additional details in supplementary statistical analysis. The results do not change noticeably when only patients from the first wave are included (supplementary Figures `r supplementary_figures_multiverse`).

@@

We also examined the association between treatments and length of hospital stay for all the patients that survived. Almost all models cannot discern the sign of the association for all treatments examined (Supplementary Figures `r supplementary_figures_multiverse`). Similarly, we studied the association between D-dimer and Interleukin 6 and outcomes, with unconclusive results as well (Supplementary Figure `r supplementary_figure_markers` )

```{r multiverseresultsdrugs, fig.height=7, fig.width=8.5, fig.cap="Estimates of model coefficients for association between treatments and main outcomes. Each row represents a model - Categorical 7/28 = Bayesian categorical regression for state at day 7/28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Figure \\@ref(fig:modelstates)a), HMM A = Bayesian hidden-Markov model as in Figure \\@ref(fig:modelstates)b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Figure \\@ref(fig:modelstates)b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Figure \\@ref(fig:modelstates)c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma.  "}

@@

Following Wynants et al.[@http://zotero.org/users/5567156/items/TS57HJ3T] we found five prediction models we were able to recompute: Li et al. report the ACP index [@http://zotero.org/users/5567156/items/EN4GICJ6] combining CRP and age to form 3 grades, Chen & Liu [@http://zotero.org/users/5567156/items/4DDUVEKS] report a continuous score using age, CRP, D-dimer and lymphocyte count,
Shi et al. [@http://zotero.org/users/5567156/items/X4GWJIP5] use age, sex and hypertension to form 4 grades, Caramelo et al. use age, sex, hypertension, diabetes, cardiac disease, chronic respiratory disease and cancer to form a continuous score [@http://zotero.org/users/5567156/items/3A2RN4ZS], Bello-Chavolla et al. [@http://zotero.org/users/5567156/items/INZVABQI] use age, diabetes, obesity, pneumonia, chronic kidney disease, COPD and immunosuppression to build a score ranging from -6 to 22. For the latter two scores we had to impute some of the predictors as they had no immediate equivalent in our dataset. The outcomes present in the studies were: 12-day mortality, 30-day mortality and mortality without any further details, here we report results for both 12-day and 30-day mortality. Full details on the scores and how we used our dataset to compute them is given in the supplementary statistical analysis.

All prognostic models we tested performed similarly to or notably worse than using age as the only predictor and also worse than originally reported (Figure&nbsp;\@ref(fig:riskscores)). Additionally, some publications did not provide enough detail to unambiguously reconstruct how the score and/or outcome was assessed. We thus concur with Wynants et al. [@http://zotero.org/users/5567156/items/TS57HJ3T] that reported prediction scores are at high risk of bias and need additional careful evaluation.

```{r riskscores, fig.cap="Performance of tested prediction scores as measured by AUC. AUC = 1 means perfect prediction while AUC $\\leq$ 0.5 means that the score is worse than random guess and a better prediction would be obtained by reversing the score (marked by thick blue line). The line ranges represent the bootstrapped 95% confidence intervals. Red dots show results computed in present study - model variants (horizontal axis) vary in the outcome measured (12-day or 30-day mortality, severe disease) and potentially on how ambiguities in score computation were resolved, although this rarely makes a big difference - see supplementary statistical analysis for details. Cyan triangles show AUC as reported by the original authors or recomputed based on their published data. When the confidence interval or the AUC of the original study is not shown, it means that the value was not reported by the authors and not enough information to recompute it was given."}
col_types_auc <- cols(
  auc = col_double(),
  auc_low = col_double(),
  auc_high = col_double(),
  outcome = col_character(),
  score = col_character(),
  note = col_character(),
  subgroup = col_character()
)
all_auc <- read_csv(here::here("manuscript", "all_auc.csv"), col_types = col_types_auc, na = "NA") %>% filter(subgroup != "(admission)")
all_auc_orig <- read_csv(here::here("manuscript", "all_auc_orig.csv"), col_types = col_types_auc, na = "NA") %>% filter(subgroup != "(admission)")

@@

>>>>

---
title: "Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in March - October 2020: An exploratory analysis" 
abstract: |  
  We collected a multi-centric retrospective dataset of patients (N = 213) who were admitted to ten hospitals in Czech Republic and tested positive for SARS-CoV-2 during the early phases of the pandemic in March - October 2020. The dataset contains baseline patient characteristics, breathing support required, pharmacological treatment received and multiple markers on daily resolution. Patients in the dataset were treated with hydroxychloroquine (N = 108), azithromycin (N = 72), favipiravir (N = 9), convalescent plasma (N = 7), dexamethasone (N = 4) and remdesivir (N = 3), often in combination. To explore association between treatments and patient outcomes we performed multiverse analysis, observing how the conclusions change between defensible choices of statistical model, predictors included in the model and other analytical degrees of freedom.  Weak evidence to constrain the potential efficacy of azithromycin and favipiravir can be extracted from the data. Additionally, we performed external validation of several proposed prognostic models for Covid-19 severity showing that they mostly perform unsatisfactorily on our dataset.
bibliography: Covid19.json
output: 
  bookdown::pdf_book:
    base_format: rticles::plos_article
    includes:
      in_header: preamble.tex

  # rticles::plos_article:
  #   includes:
  #     in_header: preamble.tex
csl: plos.csl
#date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
author: "`r { source(here::here('manuscript', 'authors.R')); get_authors_string() }`"
---

```{r setup, echo=FALSE, message=FALSE, warning = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE, fig.width=8,out.width = "100%", cache=TRUE)

@@

A convenience sample of patients from `r n_hospitals` sites was collected. The study sites span the whole spectrum of sizes from large university hospitals in major cities with multiple dedicated Covid-19 wards (Thomayer University Hospital in Prague, Motol University Hospital in Prague,  Kralovské Vinohrady University Hospital in Prague, General University Hospital in Prague, University Hospital in Pilsen) through major regional/specialized hospitals  (Na Homolce Hospital in Prague, Military Hospital Olomouc) as well as smaller hospitals caring for just several Covid-19 patients at a time (AGEL Hospital Nový Jičín, Hořovice Hospital, Třebíč Hospital). The sites were chosen based on availability and willingness of the personnel to participate in data collection. None of the study sites was exclusively dedicated to treating Covid-19 patients. For each site, the dataset contains all patients hospitalized in the participating wards over the data collection period. The data collection started at the onset of the Covid-19 pandemic in March 2020 (except for one site where some older records were inaccessible), but the end date for collection differs between sites due to time constraints of the participating physicians. Three sites included total of `r n_second_wave` patients that could be considered part of "second wave" (admitted after September 1st). Last patient included in the dataset was admitted on October 12th. See S1 Fig  for per-site data collection periods. Patients over the age of 18 were included if they had PCR-confirmed infection of SARS-CoV-2 and were not participating in a clinical trial of any Covid-19 pharmacotherapy.

@@

We collected data on comorbidities and information about disease progression on daily resolution including breathing support required, oxygen flow rate, experimental anti-Covid-19 and antimicrobial drugs taken and several laboratory markers (PCR positivity for SARS-CoV-2, C-reactive protein, D-dimer, Interleukin 6, Ferritin, lymphocyte count). Full protocol for data collection is attached in S1 file and the data collection table in S2 file. Due to very low number of patients using extra-corporeal membrane oxygenation (N = `r n_ecmo`) or non-invasive positive pressure ventilation (N = `r n_nippv`) in our sample, we merged those categories with mechanical ventilation.

@@

First class of models are frequentist survival and multistate models under the proportional hazards assumption as implemented in the `coxph` function from the `survival` package [@http://zotero.org/users/5567156/items/I9UVTUJF]. We primarily use a model with competing risks for death and discharge from hospital (see Fig&nbsp;\@ref(fig:modelstates)a).

Second class of models are Bayesian hidden Markov models (HMM) of disease progression implemented via a custom extension to the `brms` package [@http://zotero.org/users/5567156/items/7ER3ZSER]. The parametrization of the HMM is inspired by Williams et al.[@http://zotero.org/users/5567156/items/F9K2L5Q8]: the actual process of disease is assumed to be continuous and allow only for transitions between neighboring states (as shown in Fig&nbsp;\@ref(fig:modelstates)b, c). The total probability of transition between any two states over the period of a day is then computed as the total probability of transition across all possible paths. This class of models does not satisfy the proportional hazards assumption, instead, it is assumed the process has the Markov property - i.e., that the (potentially unobserved) state and the covariates at a given day contain all the information to determine probabilities of the states on the next day. We use two versions of such models, one working solely with the observed breathing support and one assuming a hidden improving/worsening distinction. All of the hidden Markov models take into account whether best supportive care was initiated and a patient was thus not indicated to progress to more intensive treatment modalities.

@@

Except for age, sex and comorbidities, all covariates are treated as time-varying, e.g., the effect of taking a drug is only included for the days after the drug was taken. More details on the exact model formulations can be found in the supplementary statistical analysis in S3 file.

@@

In total, we were able to gather data for `r n_patients` patients, see Table \@ref(tab:summarytable) for the overall characteristics of the patient sample and several subgroups we used in the analysis, including treatments taken. 
Counts of all treatment combinations are shown in S2 Fig and S3 Fig shows outcomes by study site, demonstrating quite large hospital-specific differences.
The dataset includes 19 patients already reported in a study of inflammatory signatures of Covid-19 [@http://zotero.org/users/5567156/items/IA4FQCF3].

@@

In Fig&nbsp;\@ref(fig:progression) we show the overall disease progression for all patients and in Fig&nbsp;\@ref(fig:selectedmarkers) we show the time-course of a subset of the markers we have measured. The data show some interesting patterns: patients with low Interleukin-6 or D-dimer values are overrepresented among patients with better outcomes, most patients had high CRP upon admission and for many the CRP levels stayed elevated over the whole hospitalization. However, the limited nature of the data does not allow for any statistically robust conclusions. We also see that the marker levels were not substantially stratified by study site.  Those patterns should however be interpreted with care due to systematic missingness of the data - in particular, patients that fared worse were probably more likely to have the markers measured. However, we believe this kind of patient-level view is useful to appreciate the extent of both between-patient and within-patient variability.

The between-patient variability is notable even across outcomes - when ordering the patients by the highest CRP levels experienced throughout the hospital stay, the top 20% of patients that breathed ambient air for the whole hospitalization experienced higher levels than the bottom 20% of patients that required ventilation or died. This overlap is even larger when comparing only against the patients that died and D-dimer, Interleukin-6 and lymphocyte count also show a notably larger overlap than CRP (S4 Fig).

@@

```{r selectedmarkers, fig.cap = "Values of selected markers over the course of the disease. Each line represents a patient, stratified by the worst breathing support required over the hospitalization. Color indicates study sites. The vertical scale is logarithmic. Ventilated = any form of ventilation (non-invasive positive-pressure ventilation, mechanical ventilation and extra-corporeal membrane oxygenation), CRP = C-reactive protein [mg/l], D-dimer [ng/ml DDU], Ly = lymphocyte count [$10^9$/l], IL-6 = Interleukin 6 [ng/l]."}
#data$marker_data %>%
# TODO: use SpO2 in AA as SpO2 native (and vice versa)
plot_marker_data <- function(markers, log_transform, smooth = FALSE, max_patients = NULL, max_day = NULL) {

@@

As noted above, the nature of the convenience sample did not enforce random assignment of treatments to patients. In fact, patients with worse baseline characteristics, which lead to worse outcomes, were less likely to receive hydroxychloroquine (see Fig&nbsp;\@ref(fig:hcqassoc)). This clearly creates a bias towards a positive effect of hydroxychloroquine on the outcome (and potentially for other treatments as well - most were used in combination with hydroxychloroquine), which, however, could be false.

```{r hcqassoc, message=FALSE, fig.width=7, fig.height=4.5, fig.cap="The choice of treatment with hydroxychloroquine seemed to be associated with the status of patients upon hospitalization. Comorbidities were associated with both worse outcome (black) and lower chance of treatment with hydroxychloroquine (red). Dots and lines represent the estimates and the 95\\% confidence intervals of the log odds ratio of the respective outcome. HCQ: hydroxychloroquine, IHD: ischemic heart disease, HD: hypertension drugs, HF: heart failure history, COPD: chronic obstructive pulmonary disease, LungD: other lung disease, Dia: diabetes, RenalD: renal disease, LiverD: liver disease, HighCr: creatinin above 115 for males or above 97 for females, HighInr: Prothrombin time (Quick test) as International Normalized Ratio above 1.2, LowAlb: albumin in serum/plasma below 36 g/l."}

@@

Here, we focus on hydroxychloroquine and azithromycin as those are the only treatments with larger sample size. We also investigate favipiravir as it is less well reported in the literature. Hydroxychloroquine was dosed almost exclusively in a 5-day regime starting with a loading dose of 800mg on the first day and followed by 400mg. Majority of patients complemented hydroxychloroquine with azithromycin while azithromycin was rarely used alone (see Table \@ref(tab:summarytable)). Azithromycin was most frequently dosed 250 or 500mg/day, but doses ranging from 100mg/day to 1500 mg/day were observed.  Favipiravir was used only at one site with a loading dose of 3600mg on the first day, followed by at most 9 days with a 1600mg dose. All but one of the patients receiving favipiravir also received hydroxychloroquine. Treatment was initiated mostly within two days of admission (see S5 Fig).

The results of the multiverse analysis for association between hydroxychloroquine, azithromycin and favipiravir usage and death is shown in Fig&nbsp;\@ref(fig:multiverseresultsdrugs) - here, we only show models that were not found to have immediate problems representing the data well or computational issues (see S3 file for details). Results for all models we tested are reported in  S`r supplementary_figures_multiverse` Figs, with additional details in S3 file. The results do not change noticeably when only patients from the first wave are included (S`r supplementary_figures_multiverse` Figs).

@@

We also examined the association between treatments and length of hospital stay for all the patients that survived. Almost all models cannot discern the sign of the association for all treatments examined (S`r supplementary_figures_multiverse` Figs). Similarly, we studied the association between D-dimer and Interleukin 6 and outcomes, with unconclusive results as well (S`r supplementary_figure_markers` Fig)

```{r multiverseresultsdrugs, fig.height=7, fig.width=8, fig.cap="Estimates of model coefficients for association between treatments and main outcomes. Each row represents a model - Categorical 7/28 = Bayesian categorical regression for state at day 7/28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95\\% confidence interval, for Bayesian models we show posterior mean and 95\\% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma.  "}

@@

Following Wynants et al.[@http://zotero.org/users/5567156/items/TS57HJ3T] we found five prediction models we were able to recompute: Li et al. report the ACP index [@http://zotero.org/users/5567156/items/EN4GICJ6] combining CRP and age to form 3 grades, Chen & Liu [@http://zotero.org/users/5567156/items/4DDUVEKS] report a continuous score using age, CRP, D-dimer and lymphocyte count,
Shi et al. [@http://zotero.org/users/5567156/items/X4GWJIP5] use age, sex and hypertension to form 4 grades, Caramelo et al. use age, sex, hypertension, diabetes, cardiac disease, chronic respiratory disease and cancer to form a continuous score [@http://zotero.org/users/5567156/items/3A2RN4ZS], Bello-Chavolla et al. [@http://zotero.org/users/5567156/items/INZVABQI] use age, diabetes, obesity, pneumonia, chronic kidney disease, COPD and immunosuppression to build a score ranging from -6 to 22. For the latter two scores we had to impute some of the predictors as they had no immediate equivalent in our dataset. The outcomes present in the studies were: 12-day mortality, 30-day mortality and mortality without any further details, here we report results for both 12-day and 30-day mortality. Full details on the scores and how we used our dataset to compute them is given in the S3 file.

All prognostic models we tested performed similarly to or notably worse than using age as the only predictor and also worse than originally reported (Fig&nbsp;\@ref(fig:riskscores)). Additionally, some publications did not provide enough detail to unambiguously reconstruct how the score and/or outcome was assessed. We thus concur with Wynants et al. [@http://zotero.org/users/5567156/items/TS57HJ3T] that reported prediction scores are at high risk of bias and need additional careful evaluation.

```{r riskscores, fig.cap="Performance of tested prediction scores as measured by AUC. AUC = 1 means perfect prediction while AUC $\\leq$ 0.5 means that the score is worse than random guess and a better prediction would be obtained by reversing the score (marked by thick blue line). The line ranges represent the bootstrapped 95\\% confidence intervals. Red dots show results computed in present study - model variants (horizontal axis) vary in the outcome measured (12-day or 30-day mortality, severe disease) and potentially on how ambiguities in score computation were resolved, although this rarely makes a big difference - see S3 file for details. Cyan triangles show AUC as reported by the original authors or recomputed based on their published data. When the confidence interval or the AUC of the original study is not shown, it means that the value was not reported by the authors and not enough information to recompute it was given."}
col_types_auc <- cols(
  auc = col_double(),
  auc_low = col_double(),
  auc_high = col_double(),
  outcome = col_character(),
  score = col_character(),
  note = col_character(),
  subgroup = col_character()
)
all_auc <- read_csv(here::here("manuscript", "all_auc.csv"), col_types = col_types_auc, na = "NA") %>% filter(subgroup != "(admission)")
all_auc_orig <- read_csv(here::here("manuscript", "all_auc_orig.csv"), col_types = col_types_auc, na = "NA") %>% filter(subgroup != "(admission)")

@@

# Supporting information

* **S1 Fig. Data collection timeline.** Data collection periods at individual sites, showing the range of admission dates of patients included in the study. Note that we cannot provide additional information to link the sites here with data shown elsewhere as that would  increase the risk of deanonymization of the patients.

* **S2 Fig. Treatment combinations. ** Upset plot of treatment combinations - each vertical bar displays the number of patients that received the combination indicated by filled dots in the matrix. Horizontal bars show the total number of patients receiveing the given treatment.

* **S3 Fig. Outcomes per site.** Number of patients and outcomes at the individual sites. The numbers above bars are the exact counts. Hospitalized = still hospitalized at the end of data collection at the site or transferred to other site and lost to followup. Sites are anonymized to preserve patient privacy.

* **S4 Fig. Markers and outcomes.** Density plots of worst marker values per patient, stratified by worst condition experienced by the patient. For each patient that had a given marker measured, the worst value was taken. Additionally the patients are classified by the worst condition (regardless of the timing relative to the worst marker levels). For each set of patients and marker an empirical density plot of the worst marker values is shown.

* **S5 Fig. Treatment onset.** Histogram of timing of first treatment relative to admission into one of the study sites. Two patients initiated treatment before admission, which is shown as the negative numbers.

* **S6 Fig. Association of HCQ with outcomes.** Estimates of model coefficients for association between hydroxychloroquine and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S7 Fig. Association of azithromycin with outcomes.** Estimates of model coefficients for association between azithromycin and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S8 Fig. Association of favipiravir with outcomes.** Estimates of model coefficients for association between favipiravir and main outcomes. The "Suspicious" section shows models that were found to not fit the data well or have computational issues - see supplementary statistical analysis for details.  Each row represents a model - Categorical All/7/28 = Bayesian categorical regression for state at last observed day/day 7/day 28, Binary All/7/28 = Bayesian logistic regression for state at last observed day/day 7/day 28, Bayes Cox = Bayesian version of the Cox proportional hazards model with a binary outcome, Cox (single) = frequentist Cox model with a binary outcome, Cox (competing) = frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-Markov model as in Fig 1b with predictors for rate groups, HMM B = Bayesian hidden-markov model as in Fig 1b with predictors for individual rates, HMM C = Bayesian hidden-Markov model as in Fig 1c. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. The estimands are either log odds-ratio (Categorical, HMM) or log hazard ratio (Cox variants) or log ratio of mean duration of hospitalization (HMM duration). In all cases coefficient < 0 means better patient outcome in the treatment group. Vertical lines indicate zero (blue) and substantial increase or decrease with odds or hazard ratio of 3:2 or 2:3 (green). Additionally the factors the model adjusted for are listed - Site = the study site, admitted = Admitted for Covid-19, Supportive = best supportive care initiated, Comorb. = total number of comorbidities, AZ = took azithromycin, HCQ = took hydroxychloroquine, FPV = took favipiravir, C. plasma = received convalescent plasma, first wave = only patients admitted before September 1st were included.

* **S9 Fig. Association of markers with outcomes.** Estimates of model coefficients (log hazard ratios) for association between markers and death. The "Suspicious" section shows models that were found to not fit the data well or have computational issues, "Problematic" section shows models that were completely broken - see supplementary statistical analysis for details. Each row represents a model - Cox (competing) =  frequentist Cox model using competing risks (as in Fig 1a), HMM A = Bayesian hidden-markov model as in Fig 1b with predictors for rate groups, JM = Bayesian joint longitudinal and time-to-event model. For frequentist models, we show maximum likelihood estimate and 95% confidence interval, for Bayesian models we show posterior mean and 95% credible interval. Additionally the factors the model adjusted for are listed - Site = the study site, Supportive = best supportive care initiated, HCQ = took Hydroxychloroquine. We show posterior mean and 95% credible interval.

* **S1 File. Data collection protocol.**

* **S2 File. MS Excel table used for data collection.**
